### 2019-11-15 thesis/content/chapter-related-work.tex
started with gcr related work

Before combining \ac{lta} and \ac{gcr} as described in \cref{sec:intro:goals}, we first give an overview of the state-of-the-art in both fields of research.
First the existing \ac{lta} methods for unstructured set inputs are described.
Then we will look at the two main families of \ac{gcr} methods:
\begin{enumerate*}
	\item Vector representation methods
	\item Graph neural networks
\end{enumerate*}.

@@

The first approach uses \textit{uninorms}~\cite{Melnikov2016} to do so.
There the basic idea is to express composite scores as fuzzy truth assignments $y \in [0, 1]$.
Such a composite assignment $y$ is modeled as the result of a parameterized logical expression of constituent assignments $y_i \in [0, 1]$.
As the logical expression that thus effectively aggregates the constituents, a uninorm $U_{\lambda}$ is used.
Depending on the parameter $\lambda \in [0, 1]$, $U_{\lambda}$ combines a t-norm $T$ and a t-conorm $S$ which are continuous generalizations of logical conjunction and disjunction respectively.
One popular choice of norms are the so-called Łukasiewicz norms:
\begin{align}
	\text{t-norm } T(a, b) &:= \max \{ 0, a + b - 1 \}, \quad\text{t-conorm } S(a, b) := \min \{ a + b, 1 \} \nonumber \\
	\text{uninorm } U_\lambda(a, b) &:= \begin{cases}
		\lambda T\left(\frac{a}{\lambda}, \frac{b}{\lambda}\right) & \text{if } a, b \in [0, \lambda] \\
		\lambda + (1 - \lambda) S\left(\frac{a - \lambda}{1 - \lambda}, \frac{b - \lambda}{1 - \lambda}\right) & \text{if } a, b \in [\lambda, 1] \\
		\lambda \min \{ a, b \} & \text{else}
	\end{cases}
\end{align}
At the extreme points $(0, 0)$, $(0, 1)$, $(1, 0)$ and $(1, 1)$, $T$ and $S$ coincide with the Boolean operators $\land$ and $\lor$;
the values at all other points are interpolated as shown in \cref{fig:related:logic-norms}.
The uninorm $U_\lambda$ uses the conjunctive t-norm $T$ for values below the threshold $\lambda$ and the disjunctive t-conorm $S$ for values above the threshold.
$U_\lambda$ therefore smoothly interpolates between a conjunctive and disjunctive operator with the extreme points $U_1 = T$ and $U_0 = S$.
\begin{figure}[h]
	\centering
	\begin{tikzpicture}
		\begin{axis}[
			title={t-norm $T$ ($\land$)},
			xlabel=$a$,
			ylabel=$b$,
			xlabel style={xshift=0.2cm, yshift=0.15cm},
			ylabel style={xshift=-0.2cm, yshift=0.15cm},
			tick label style={font=\scriptsize},
			width=0.33\textwidth,
			colormap = {bluered}{color(0cm) = (t_blue); color(1cm) = (t_red)}
		]
			\addplot3[
				mesh,
				samples=12,
				domain=0:1,
				domain y=0:1
			]{max(x + y - 1, 0)};
		\end{axis}
	\end{tikzpicture}
	\begin{tikzpicture}
		\begin{axis}[
			title={t-conorm $S$ ($\lor$)},
			xlabel=$a$,
			ylabel=$b$,
			xlabel style={xshift=0.2cm, yshift=0.15cm},
			ylabel style={xshift=-0.2cm, yshift=0.15cm},
			tick label style={font=\scriptsize},
			width=0.33\textwidth,
			colormap = {bluered}{color(0cm) = (t_blue); color(1cm) = (t_red)}
		]
			\addplot3[
				mesh,
				samples=12,
				domain=0:1,
				domain y=0:1
			]{min(x + y, 1)};
		\end{axis}
	\end{tikzpicture}
	\begin{tikzpicture}
		\begin{axis}[
			title={uninorm $U_{0.5}$ ($\land$/$\lor$)},
			xlabel=$a$,
			ylabel=$b$,
			xlabel style={xshift=0.2cm, yshift=0.15cm},
			ylabel style={xshift=-0.2cm, yshift=0.15cm},
			tick label style={font=\scriptsize},
			width=0.33\textwidth,
			colormap = {bluered}{color(0cm) = (t_blue); color(1cm) = (t_red)}
		]
			\addplot3[
				mesh,
				samples=12,
				domain=0:1,
				domain y=0:1
			]{(x <= 0.5 && y <= 0.5) * 0.5 * max(2 * (x + y) - 1, 0) + (x > 0.5 && y > 0.5) * (0.5 + 0.5 * min(2 * (x + y - 1), 1)) + (!(x <= 0.5 && y <= 0.5) && !(x > 0.5 && y > 0.5)) * min(x, y)};
		\end{axis}
	\end{tikzpicture}
	\caption{The Łukasiewicz norms and the corresponding uninorm for $\lambda = 0.5$.}\label{fig:related:logic-norms}
\end{figure}

@@

Recently \citet{Melnikov2019} have looked at an alternative class of aggregation functions.
Instead of using fuzzy logic to describe score aggregation, \ac{owa} operators were used.
\ac{owa} aggregators work by sorting the input scores and then weighting them based on their sort position, i.e.\ %
\begin{align*}
	\mathcal{A}_{\lambda}(y_1, \dots, y_n) := \sum_{i = 1}^n \lambda_i y_{\pi(i)},
\end{align*}
where $\lambda \in \mathbb{R}^n$ is a weight vector with ${\|\lambda\|}_1 = 1$ and $\pi$ is a sorting permutation of the input scores with $y_i < y_j \Rightarrow \pi(i) < \pi(j)$. % chktex 21
Depending on the choice of the vector $\lambda$, the \ac{owa} function $\mathcal{A}_\lambda$ can express common aggregation functions like
$\min$ (if $\lambda = (1, 0, \dots, 0)$),
$\max$ (if $\lambda = (0, \dots, 0, 1)$)
or the arithmetic mean (if $\lambda = \left(\frac{1}{n}, \dots, \frac{1}{n}\right)$).

@@

\begin{figure}[h]
	\centering
	\includegraphics[width=0.75\linewidth]{gfx/related-work/bum.pdf}
	\caption[Illustration of how a \ac{bum} function is described as a linear spline and its relation to the \ac{owa} weights.]{
		Illustration of how \textcolor{t_blue}{$a$} describes \textcolor{t_blue}{$q$} and its relation to \textcolor{t_red}{$\lambda$} ($n = 4$, $m = 7$).
	}\label{fig:related:bum}
\end{figure}
In the \ac{owa} approach the \ac{bum} function $q$ is modeled as a piecewise linear spline.
This spline is described by $m+1$ points ${\left\{ \left( \frac{j}{m}, a_j \right) \right\}}_{j = 0}^{m}$, the so-called knots of the spline. % chktex 21
The curve of $q$ is obtained by linearly interpolating between neighboring knots as shown in \cref{fig:related:bum}.
If $0 = a_0 \leq a_1 \leq \cdots \leq a_m = 1$, $q$ is a \ac{bum} function.
The \ac{lta} aggregation problem is therefore solved by optimizing $a \in \mathbb{R}^{m + 1}$ under this constraint.
The disaggregation problem is tackled by adding the scores $(y_1, \dots, y_M) \in \mathbb{R}^M$ to the learnable parameters of the model where $M$ is assumed to be the finite number of constituents.
Currently the \ac{owa} approach requires all possible constituents to be part of the training dataset since it does not consider constituent features $x_i$ to predict the scores of unseen constituents.

@@

\subsection{Graph Embeddings}%
\label{sec:related:gcr:embed}

>>>>

Before combining \ac{lta} and \ac{gcr} as described in \cref{sec:intro:goals}, we first give an overview of the state-of-the-art in both fields of research.
First the existing \ac{lta} methods for unstructured set inputs are described.
Then we will look at the current approaches to tackle the \ac{gcr} problem.

@@

The first approach uses \textit{uninorms}~\cite{Melnikov2016} to do so.
There the basic idea is to express composite scores as fuzzy truth assignments $y \in [0, 1]$.
Such a composite assignment $y$ is modeled as the result of a parameterized logical expression of constituent assignments $y_i \in [0, 1]$.
As the logical expression that thus effectively aggregates the constituents, a uninorm $U_{\lambda}$ is used.
Depending on the parameter $\lambda \in [0, 1]$, $U_{\lambda}$ combines a t-norm $T$ and a t-conorm $S$ which are continuous generalizations of logical conjunction and disjunction respectively.
One popular choice of norms are the so-called Łukasiewicz norms:
\begin{align}
	\text{t-norm } T(a, b) &:= \max \{ 0, a + b - 1 \}, \quad\text{t-conorm } S(a, b) := \min \{ a + b, 1 \}, \nonumber \\
	\text{uninorm } U_\lambda(a, b) &:= \begin{cases}
		\lambda T\left(\frac{a}{\lambda}, \frac{b}{\lambda}\right) & \text{if } a, b \in [0, \lambda] \\
		\lambda + (1 - \lambda) S\left(\frac{a - \lambda}{1 - \lambda}, \frac{b - \lambda}{1 - \lambda}\right) & \text{if } a, b \in [\lambda, 1] \\
		\lambda \min \{ a, b \} & \text{else}
	\end{cases}
\end{align}
At the extreme points $(0, 0)$, $(0, 1)$, $(1, 0)$ and $(1, 1)$, $T$ and $S$ coincide with the Boolean operators $\land$ and $\lor$;
the values at all other points are interpolated as shown in \cref{fig:related:logic-norms}.
The uninorm $U_\lambda$ uses the conjunctive t-norm $T$ for values below the threshold $\lambda$ and the disjunctive t-conorm $S$ for values above the threshold.
$U_\lambda$ therefore smoothly interpolates between a conjunctive and disjunctive operator with the extreme points $U_1 = T$ and $U_0 = S$.
\begin{figure}[h]
	\centering
	\begin{tikzpicture}
		\begin{axis}[
			title={t-norm $T$ ($\land$)},
			xlabel=$a$,
			ylabel=$b$,
			xlabel style={xshift=0.2cm, yshift=0.15cm},
			ylabel style={xshift=-0.2cm, yshift=0.15cm},
			tick label style={font=\scriptsize},
			width=0.33\textwidth,
			colormap = {bluered}{color(0cm) = (t_blue); color(1cm) = (t_red)}
		]
			\addplot3[
				mesh,
				samples=12,
				domain=0:1,
				domain y=0:1
			]{max(x + y - 1, 0)};
		\end{axis}
	\end{tikzpicture}
	\begin{tikzpicture}
		\begin{axis}[
			title={t-conorm $S$ ($\lor$)},
			xlabel=$a$,
			ylabel=$b$,
			xlabel style={xshift=0.2cm, yshift=0.15cm},
			ylabel style={xshift=-0.2cm, yshift=0.15cm},
			tick label style={font=\scriptsize},
			width=0.33\textwidth,
			colormap = {bluered}{color(0cm) = (t_blue); color(1cm) = (t_red)}
		]
			\addplot3[
				mesh,
				samples=12,
				domain=0:1,
				domain y=0:1
			]{min(x + y, 1)};
		\end{axis}
	\end{tikzpicture}
	\begin{tikzpicture}
		\begin{axis}[
			title={uninorm $U_{0.5}$ ($\land$/$\lor$)},
			xlabel=$a$,
			ylabel=$b$,
			xlabel style={xshift=0.2cm, yshift=0.15cm},
			ylabel style={xshift=-0.2cm, yshift=0.15cm},
			tick label style={font=\scriptsize},
			width=0.33\textwidth,
			colormap = {bluered}{color(0cm) = (t_blue); color(1cm) = (t_red)}
		]
			\addplot3[
				mesh,
				samples=12,
				domain=0:1,
				domain y=0:1
			]{(x <= 0.5 && y <= 0.5) * 0.5 * max(2 * (x + y) - 1, 0) + (x > 0.5 && y > 0.5) * (0.5 + 0.5 * min(2 * (x + y - 1), 1)) + (!(x <= 0.5 && y <= 0.5) && !(x > 0.5 && y > 0.5)) * min(x, y)};
		\end{axis}
	\end{tikzpicture}
	\caption{The Łukasiewicz norms and the corresponding uninorm for $\lambda = 0.5$.}\label{fig:related:logic-norms}
\end{figure}

@@

Recently \citet{Melnikov2019} have looked at an alternative class of aggregation functions.
Instead of using fuzzy logic to describe score aggregation, \ac{owa} operators were used.
\ac{owa} aggregators work by sorting the input scores and then weighting them based on their sort position, i.e.\ %
\begin{align}
	\mathcal{A}_{\lambda}(y_1, \dots, y_n) := \sum_{i = 1}^n \lambda_i y_{\pi(i)},
\end{align}
where $\lambda \in \mathbb{R}^n$ is a weight vector with ${\|\lambda\|}_1 = 1$ and $\pi$ is a sorting permutation of the input scores with $y_i < y_j \Rightarrow \pi(i) < \pi(j)$. % chktex 21
Depending on the choice of the vector $\lambda$, the \ac{owa} function $\mathcal{A}_\lambda$ can express common aggregation functions like
$\min$ (if $\lambda = (1, 0, \dots, 0)$),
$\max$ (if $\lambda = (0, \dots, 0, 1)$)
or the arithmetic mean (if $\lambda = \left(\frac{1}{n}, \dots, \frac{1}{n}\right)$).

@@

\begin{figure}[h]
	\centering
	\includegraphics[width=0.75\linewidth]{gfx/related-work/bum.pdf}
	\caption[Illustration of how a \ac{bum} function is described as a linear spline and its relation to the \ac{owa} weights.]{
		Illustration of how \textcolor{t_blue}{$a$} describes \textcolor{t_blue}{$q$} and its relation to \textcolor{t_red}{$\lambda$} ($n = 4$, $m = 7$).
	}\label{fig:related:bum}
\end{figure}
In the \ac{owa} approach the \ac{bum} function $q$ is modeled as a piecewise linear spline.
This spline is described by $m+1$ points ${\left\{ \left( \frac{j}{m}, a_j \right) \right\}}_{j = 0}^{m}$, the so-called knots of the spline. % chktex 21
The curve of $q$ is obtained by linearly interpolating between neighboring knots as shown in \cref{fig:related:bum}.
If $0 = a_0 \leq a_1 \leq \cdots \leq a_m = 1$, $q$ is a \ac{bum} function.
The \ac{lta} aggregation problem is therefore solved by optimizing $a \in \mathbb{R}^{m + 1}$ under this constraint.
The disaggregation problem is tackled by adding the scores $y_1, \dots, y_M \in \mathbb{R}$ to the learnable parameters of the model where $M$ is assumed to be the finite number of constituents.
Currently the \ac{owa} approach requires all possible constituents to be part of the training dataset since it does not consider constituent features $x_i \in \mathcal{X}$ to predict the scores of previously unseen constituents.

@@

The existing approaches to tackle the \ac{gcr} problem can be categorized into three main families:
\begin{enumerate*}
	\item Explicit graph embeddings,
	\item Graph kernels and
	\item Graph neural networks
\end{enumerate*}.
We will now look at the characteristics of those families and give an overview of specific methods.

\subsection{Explicit Graph Embeddings}%
\label{sec:related:gcr:embed}

The basic idea of graph embedding approaches is to map a graph $G$ to some finite vector space $\mathbb{R}^d$.
By embedding a graph any classification or regression algorithm that works with vectors can then be applied.
There are three main types of graph embedding approaches.

\paragraph{Fingerprint embeddings}

\paragraph{Unsupervised embeddings}

\paragraph{Supervised embeddings}
